An assessment of photometric redshift PDF performance in the context of LSST

### 2018-09-27 paper/tex/introduction.tex
start of introduction edits

Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.

@@

The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
In addition to the contribution of low signal-to-noise photometry to the systematic error of photo-$z$s, these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in previous large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.

There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.  However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.  For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.  A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.  Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift. %\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
As more groups began to employ photometric redshifts in their cosmological analyses, realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.

@@

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.
This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.

Certain science cases need redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases need only ensemble redshift information; for instance current weak lensing techniques require the overall redshift distribution $N(z)$ for tomographic redshift samples, but do not need single galaxy estimates.
In the case of the multiple types of probes of cosmology enabled by the
LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that meet the needs of all such science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.

>>>>

Large-scale photometric galaxy surveys are entering a new era with currently or soon-to-be running Stage III and Stage IV dark energy experiments like the Dark Energy Survey \citep[DES,][]{Abbott:05}, the Kilo-Degree Survey \citep[KiDS,][]{de_Jong:13}, Hyper Suprime-Cam (HSC) Survey \citep[]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[LSST,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[WFIRST,][]{Green:12}.
The move to imaging based surveys, rather than spectroscopic based, for cosmological measurements makes proper understanding of photometric redshifts (``photo-$z$'s'') of paramount importance, as cosmological distance measures for statistical samples are directly dependent on photo-$z$ measurements.

For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases seem to need only ensemble redshift information; for instance current weak lensing techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.  However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.  These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.  Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
%In the case of the multiple types of probes of cosmology enabled by the
%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalog compression schemes that balance information retention against limited storage resources.

@@

The tremendous size of LSST's galaxy catalog will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.
In order to meet LSST's demanding error budget, it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood.

There is often a desire to have a single valued ``point-estimate'' redshift for an individual galaxy.  However, the complex, non-linear (and often non-unique) nature of the mapping between broad band fluxes and redshift means that a single value is unable to capture the full redshift information encoded in a galaxy's magnitudes.  For example, a common point-estimate for a template-based method is taking the highest likelihood solution as the point photo-$z$.  A single valued redshift ignores degenerate redshift solutions of lower probability, potentially biasing photometric redshift estimates both for individual galaxies and ensemble distributions.  Storing more information is necessary, most often photo-$z$ codes output the redshift probability density function (PDF), also often referred to as $p(z)$, describing the relative likelihood as a function of redshift. %\red{should we move some of AIM's p(z) vs p(z|d,I) to here?}
Early template methods such as \citet{Fernandezsoto:99} converted relative $\chi^2$ values of template spectra to likelihoods to estimate $p(z)$.
Soon after, codes such as \citet{Benitez:00} added a Bayesian prior and output a posterior probability distribution.
While many early machine learning based algorithms focused on a point-estimate, \citet{Firth:03} used a neural net with 1000 realizations scattered within the photometric errors to estimate a $p(z)$.
As more groups began to employ photometric redshifts in their cosmological analyses, there was a realization that point-estimate photo-$z$'s were inadequate for precision cosmology measurements \citep{Mandelbaum:2008}.
From around this point onward, most photo-$z$ algorithms have attempted to implement some estimate of the overall redshift probability in their outputs, and some surveys began supplying a full $p(z)$ rather than a simple redshift point-estimate and error \citep[e.~g.~][]{de_Jong:17}.

@@

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
This paper is distinguished by its inclusion of metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST Dark Energy Science Collaboration (LSST-DESC).
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and their ability to produce accurate PDFs in the presence of representative training sets.
This can be thought of as an initial test under near perfect conditions, before further complexities are added in future papers.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.
